---
title: "830-pj-phase2"
author: "Yuqing Zhang"
date: "2021/8/15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
blue_palette <- colorRampPalette(
  c(rgb(247,251,255,maxColorValue = 255), 
    rgb(8,48,107,maxColorValue = 255)))
# Function for converting from natural units to coded units
convert.N.to.C <- function(U,UH,UL){
  x <- (U - (UH+UL)/2) / ((UH-UL)/2)
  return(x)
}

# Function for converting from coded units to natural units
convert.C.to.N <- function(x,UH,UL){
  U <- x*((UH-UL)/2) + (UH+UL)/2
  return(U)
}

# Function to create x and y grids for contour plots 
mesh <- function(x, y) { 
  Nx <- length(x)
  Ny <- length(y)
  list(
    x = matrix(nrow = Nx, ncol = Ny, data = x),
    y = matrix(nrow = Nx, ncol = Ny, data = y, byrow = TRUE)
  )
}
```

```{r}
data_p2 <- read.csv("../data/P2-phase1.csv")
table(data_p2$Prev.Length, data_p2$Match.Score, data_p2$Prev.Type)
data_p2_TT <- data_p2[data_p2$Prev.Type == "TT",]
data_p2_AC <- data_p2[data_p2$Prev.Type == "AC",]

data_p2_AC
```

X1: Prev.Length 100 120

X2: Match.Score 80 100

```{r}
coded_data_p2_TT <- data.frame(y = data_p2_TT$Browse.Time,
                  x1 = convert.N.to.C(U = data_p2_TT$Prev.Length, 
                                      UH = 120, UL = 100),
                  x2 = convert.N.to.C(U = data_p2_TT$Match.Score, 
                                      UH = 100, UL = 80))
coded_data_p2_TT$xPQ <- (coded_data_p2_TT$x1^2 + coded_data_p2_TT$x2^2)/2

aggregate(coded_data_p2_TT$y, 
          by = list(x1 = coded_data_p2_TT$x1, x2 = coded_data_p2_TT$x2), 
          FUN = mean)

```

```{r}
mean(coded_data_p2_TT$y[coded_data_p2_TT$xPQ != 0]) - mean(coded_data_p2_TT$y[coded_data_p2_TT$xPQ == 0])
```

```{r}
m_TT <- lm(y~x1+x2+x1*x2+xPQ, data = coded_data_p2_TT)
summary(m_TT)
```

```{r}
m.TT.fo <- lm(y~x1+x2, data = coded_data_p2_TT)
beta0 <- coef(m.TT.fo)[1]
beta1 <- coef(m.TT.fo)[2]
beta2 <- coef(m.TT.fo)[3]
grd <- mesh(x = seq(convert.N.to.C(U = 30, UH = 120, UL = 100), 
                    convert.N.to.C(U = 120, UH = 120, UL = 100), 
                    length.out = 100), 
            y = seq(convert.N.to.C(U = 0, UH = 100, UL = 80), 
                    convert.N.to.C(U = 100, UH = 100, UL = 80), 
                    length.out = 100))
x1 <- grd$x
x2 <- grd$y
eta.fo <- beta0 + beta1*x1 + beta2*x2
# 2D contour plot
contour(x = seq(convert.N.to.C(U = 30, UH = 120, UL = 100), 
                convert.N.to.C(U = 120, UH = 120, UL = 100), 
                length.out = 100),
        y = seq(convert.N.to.C(U = 0, UH = 100, UL = 80), 
                convert.N.to.C(U = 100, UH = 100, UL = 80), 
                length.out = 100), 
        z = eta.fo, xlab = "x1 (Preview Length)", ylab = "x2 (Match.Score)",
        nlevels = 15, col = blue_palette(15), labcex = 0.9, asp=1)

abline(a = 0, b = beta2/beta1, lty = 2)
points(x = 0, y = 0, col = "red", pch = 16)
```

```{r}
# The gradient vector
g <- matrix(c(beta1, beta2), nrow = 1)
g
# We will take steps of size 5 seconds in preview length. In coded units this is
PL.step <- convert.N.to.C(U = 110 + 5, UH = 120, UL = 100)
lambda <- PL.step/abs(beta1)
```


```{r}
contour(x = seq(convert.N.to.C(U = 30, UH = 120, UL = 100), 
                convert.N.to.C(U = 120, UH = 120, UL = 100), 
                length.out = 100),
        y = seq(convert.N.to.C(U = 0, UH = 100, UL = 80), 
                convert.N.to.C(U = 100, UH = 100, UL = 80), 
                length.out = 100), 
        z = eta.fo, xlab = "x1 (Preview Length)", ylab = "x2 (Match.Score)",
        nlevels = 15, col = blue_palette(15), labcex = 0.9, asp=1)

abline(a = 0, b = beta2/beta1, lty = 2)
points(x = 0, y = 0, col = "red", pch = 16)

## Step 0: The center point we've already observed
x.old <- matrix(0, nrow=1, ncol=2)
text(x = 0, y = 0+0.25, labels = "0")
step0 <- data.frame(Prev.Length = convert.C.to.N(x = 0, UH = 120, UL = 100), 
                 Match.Score = convert.C.to.N(x = 0, UH = 100, UL = 80))

## Step 1: 
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "1")
step1 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 2: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "2")
step2 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 3: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "3")
step3 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 4: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "4")
step4 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 5: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "5")
step5 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 6: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "6")
step6 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 7:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "7")
step7 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 8:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "8")
step8 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 9:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "9")
step9 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 10:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "10")
step10 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 11:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "11")
step11 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 12:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "12")
step12 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## Step 13:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "13")
step13 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 120, UL = 100), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 100, UL = 80))

## The following is a list of the conditions along the path of steepest descent
pstd.cond <- data.frame(Step = 0:13, rbind(step0, step1, step2, step3, step4, step5, step6, 
                                            step7, step8, step9, step10, step11, step12, step13))
pstd.cond
```

```{r}
pstd.cond$Match.Score <- round(pstd.cond$Match.Score)
pstd.cond
```

```{r}
pstd <- read.csv("../data/P2-phase2.csv", header = TRUE)
pstd.TT <- pstd[pstd$Prev.Type == "TT",]
pstd.means <- aggregate(pstd.TT$Browse.Time, 
                        by = list(Prev.Length = pstd.TT$Prev.Length, 
                                  Prev.Size = pstd.TT$Match.Score), 
                        FUN = mean)

plot(x = 0:3, y = pstd.means$x[1:4],
     type = "l", xlab = "Step Number", ylab = "Average Browsing Time")
points(x = 0:3, y = pstd.means$x[1:4],
       col = "red", pch = 16)
```
100, 81
105,95 -- 86,76



```{r}
head(pstd.cond)
```

```{r}
## Load this data and check whether the pure quadratic effect is significant
P2_PH3_TT <- read.csv("../data/P2-phase3-TT.csv", header = TRUE)
P2_PH3_TT_2 <- read.csv("../data/P2-phase3-TT-2.csv", header = TRUE)
temp <- rbind(P2_PH3_TT, P2_PH3_TT_2)
P2_PH3_TT <- temp[temp$Match.Score != 85, ]

P2_PH3_TT
ph3.TT <- data.frame(y = P2_PH3_TT$Browse.Time,
                  x1 = convert.N.to.C(U = P2_PH3_TT$Prev.Length, UH = 105, UL = 95),
                  x2 = convert.N.to.C(U = P2_PH3_TT$Match.Score, UH = 86, UL = 76))
ph3.TT$xPQ <- (ph3.TT$x1^2 + ph3.TT$x2^2)/2

## Check the average browsing time in each condition:
aggregate(ph3.TT$y, by = list(x1 = ph3.TT$x1, x2 = ph3.TT$x2), FUN = mean)

## The difference in average browsing time in factorial conditions vs. the center 
## point condition
mean(ph3.TT$y[ph3.TT$xPQ != 0]) - mean(ph3.TT$y[ph3.TT$xPQ == 0])

## Check to see if that's significant
m <- lm(y~x1+x2+x1*x2+xPQ, data = ph3.TT)
summary(m)
```

# AGAIN

```{r}
aggregate(ph3.TT$y, 
          by = list(x1 = ph3.TT$x1, x2 = ph3.TT$x2), 
          FUN = mean)

```

```{r}
mean(ph3.TT$y[ph3.TT$xPQ != 0]) - mean(ph3.TT$y[ph3.TT$xPQ == 0])
```

```{r}
m_TT <- lm(y~x1+x2+x1*x2+xPQ, data = ph3.TT)
summary(m_TT)
```

```{r}
m.TT.fo <- lm(y~x1+x2, data = ph3.TT)
beta0 <- coef(m.TT.fo)[1]
beta1 <- coef(m.TT.fo)[2]
beta2 <- coef(m.TT.fo)[3]
grd <- mesh(x = seq(convert.N.to.C(U = 50, UH = 105, UL = 95), 
                    convert.N.to.C(U = 120, UH = 105, UL = 95), 
                    length.out = 100), 
            y = seq(convert.N.to.C(U = 50, UH = 86, UL = 76), 
                    convert.N.to.C(U = 100, UH = 86, UL = 76), 
                    length.out = 100))
x1 <- grd$x
x2 <- grd$y
eta.fo <- beta0 + beta1*x1 + beta2*x2
# 2D contour plot
contour(x = seq(convert.N.to.C(U = 50, UH = 105, UL = 95), 
                    convert.N.to.C(U = 120, UH = 105, UL = 95), 
                    length.out = 100), 
            y = seq(convert.N.to.C(U = 50, UH = 86, UL = 76), 
                    convert.N.to.C(U = 100, UH = 86, UL = 76), 
                    length.out = 100), 
        z = eta.fo, xlab = "x1 (Preview Length)", ylab = "x2 (Match.Score)",
        nlevels = 15, col = blue_palette(15), labcex = 0.9, asp=1)

abline(a = 0, b = beta2/beta1, lty = 2)
points(x = 0, y = 0, col = "red", pch = 16)
```

```{r}
# The gradient vector
g <- matrix(c(beta1, beta2), nrow = 1)

# We will take steps of size 5 seconds in preview length. In coded units this is
PL.step <- convert.N.to.C(U = 105, UH = 105, UL = 95)
lambda <- PL.step/abs(beta1)
```


```{r}
contour(x = seq(convert.N.to.C(U = 50, UH = 105, UL = 95), 
                    convert.N.to.C(U = 120, UH = 105, UL = 95), 
                    length.out = 100), 
            y = seq(convert.N.to.C(U = 50, UH = 86, UL = 76), 
                    convert.N.to.C(U = 100, UH = 86, UL = 76), 
                    length.out = 100), 
        z = eta.fo, xlab = "x1 (Preview Length)", ylab = "x2 (Match.Score)",
        nlevels = 15, col = blue_palette(15), labcex = 0.9, asp=1)

abline(a = 0, b = beta2/beta1, lty = 2)
points(x = 0, y = 0, col = "red", pch = 16)

## Step 0: The center point we've already observed
x.old <- matrix(0, nrow=1, ncol=2)
text(x = 0, y = 0+0.25, labels = "0")
step0 <- data.frame(Prev.Length = convert.C.to.N(x = 0, UH = 105, UL = 95), 
                 Match.Score = convert.C.to.N(x = 0, UH = 86, UL = 76))

## Step 1: 
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "1")
step1 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 105, UL = 95), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 86, UL = 76))

## Step 2: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "2")
step2 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 105, UL = 95), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 86, UL = 76))

## Step 3: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "3")
step3 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 105, UL = 95), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 86, UL = 76))

## Step 4: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "4")
step4 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 105, UL = 95), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 86, UL = 76))

## Step 5: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "5")
step5 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 105, UL = 95), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 86, UL = 76))

## Step 6: 
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "6")
step6 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 105, UL = 95), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 86, UL = 76))

## Step 7:
x.old <- x.new
x.new <- x.old - lambda*g
points(x = x.new[1,1], y = x.new[1,2], col = "red", pch = 16)
text(x = x.new[1,1], y = x.new[1,2]+0.25, labels = "7")
step7 <- data.frame(Prev.Length = convert.C.to.N(x = x.new[1,1], UH = 105, UL = 95), 
                    Match.Score = convert.C.to.N(x = x.new[1,2], UH = 86, UL = 76))

## The following is a list of the conditions along the path of steepest descent
pstd.cond <- data.frame(Step = 0:7, rbind(step0, step1, step2, step3, step4, step5, step6, 
                                            step7))
pstd.cond$Match.Score <- round(pstd.cond$Match.Score)
pstd.cond
```

```{r}
pstd.means.old <- pstd.means
pstd <- read.csv("../data/P2-phase4-TT.csv", header = TRUE)
pstd.TT <- pstd[pstd$Prev.Type == "TT",]
pstd.means <- aggregate(pstd.TT$Browse.Time, 
                        by = list(Prev.Length = pstd.TT$Prev.Length, 
                                  Prev.Size = pstd.TT$Match.Score), 
                        FUN = mean)

par(mfrow = c(1,2))
plot(x = 0:3, y = pstd.means.old$x[1:4],
     type = "l", xlab = "Step Number", ylab = "Average Browsing Time")
points(x = 0:3, y = pstd.means.old$x[1:4],
       col = "red", pch = 16)

plot(x = 0:4, y = pstd.means$x[1:5],
     type = "l", xlab = "Step Number", ylab = "Average Browsing Time")
points(x = 0:4, y = pstd.means$x[1:5],
       col = "red", pch = 16)


```

```{r}
pstd.cond[pstd.cond$Step == 3,]
pstd.cond
```

85--72
95,75 -- 77,67

first time:
Prev.Length	Match.Score	Prev.Type
90	75	TT
80	75	TT
90	69	TT
80	69	TT
85	72	TT

second time:
Prev.Length	Match.Score	Prev.Type
95	77	TT
75	77	TT
95	67	TT
75	67	TT
85	72	TT

Prev.Length	Match.Score	Prev.Type
75	62	TT
75	82	TT
95	62	TT
95	82	TT
85	72	TT



```{r}
## Load this data and check whether the pure quadratic effect is significant
P2_PH5_TT <- read.csv("../data/P2-phase5.555-TT.csv", header = TRUE)

head(P2_PH5_TT)

ph5.TT <- data.frame(y = P2_PH5_TT$Browse.Time,
                  x1 = convert.N.to.C(U = P2_PH5_TT$Prev.Length, UH = 95, UL = 75),
                  x2 = convert.N.to.C(U = P2_PH5_TT$Match.Score, UH = 82, UL = 62))

ph5.TT$xPQ <- (ph5.TT$x1^2 + ph5.TT$x2^2)/2

## Check the average browsing time in each condition:
aggregate(ph5.TT$y, by = list(x1 = ph5.TT$x1, x2 = ph5.TT$x2), FUN = mean)

## The difference in average browsing time in factorial conditions vs. the center 
## point condition
mean(ph5.TT$y[ph5.TT$xPQ != 0]) - mean(ph5.TT$y[ph5.TT$xPQ == 0])

## Check to see if that's significant
m <- lm(y~x1+x2+x1*x2+xPQ, data = ph5.TT)
summary(m)
```

In each condition, we used the Method of Steepest Descent. We used a $2^2$ factorial experiment to estimate a first-order response surface: 
$$
\widehat{\eta}=\widehat{\beta}_{0}+\widehat{\beta}_{1} x_{1}+\widehat{\beta}_{2} x_{2}
$$

The gradient of this surface is then calculated: 
$$
\mathrm{g}=\nabla \widehat{\eta}=\left[\begin{array}{cccc}
\frac{\partial \hat{\eta}}{\partial x_{1}} & \frac{\partial \hat{\eta}}{\partial x_{2}}
\end{array}\right]^{T}
$$
